Modelling quasar outflows: clumpy winds and broad emission lines

### 2015-07-23 paper.tex
finished for the day

%% Journal definitions
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\apjs{ApJs}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}

@@

\title
{
The effect of clumpy outflows
and disc anisotropy on quasar unification scenarios
}

@@

The spectral energy distributions (SEDs) of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is normally attributed to emission from a geometrically thin, 
optically thick accretion disc surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material, 
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
natural feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required by models of galaxy evolution (REFs) 
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

Approximately $20\%$ of quasars exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.
In addition, ADWs may offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO, LoBAL-QSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) is normally
assumed to correspond to the dense wind base.
% As well as acting as a source of photoionized plasma, a 
A disc wind may also have a profound effect on the structure and 
emergent continuum of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN \citep{laordavis2014}.
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN SEDs can, in general, be fitted well with accretion disc models.
Mass-loss therefore appears to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

@@

The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma. \cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.

An alternative solution is that the wind is clumped (REFs), possibly on multiple scale lengths.
Local density enhancements could lower the ionization parameter of the plasma
while still maintaining the same mass-loss rate and column density. 
Clumping is expected in outflows and the BLR in a number of separate models
with different motivations.

Physically, clumping can be present due to magnetic confinement of 
dense regions (REFs) or the instabilities that are theoretically predicted 
in line-driven winds (REFs).

@@

In this paper, we attempt to test the disc wind paradigm 
using Monte Carlo radiative transfer and photoionization calculations.
In section 2, we describe our code. In section 3, we briefly discuss some 
of the successes and failures of our previous benchmark model for BALQSOs 
\citep[][hereafter H13]{higginbottom2013} and outline the model, including 
a description of our clumping prescription. In section 4, we present the results 
from a clumped model which successfully reproduces
the observed ionization state while maintaing realistic X-ray properties.
In section 5 we discuss our results, focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta interacting with either `simple ions'
or `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
The scheme allows one to treat non-LTE line transfer without
approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions. In this study,
only H \& He are treated as macro-atoms, because 
we expect recombination to be important
in determining their level populations (see M15).

@@

To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
we model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of bands is modeled as a normalised power law 
\begin{equation}
J_{\nu,j}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
or an exponential 
\begin{equation}
J_{\nu,j}=K_{exp}e^{(-h\nu/kT_{exp})}.
\end{equation}
The photoionization rate from ion $i$ can then be written as

\begin{equation}
R_{i,i+1}(J)= 
\displaystyle{n_i \sum_{band~j=0}^{n}}~{\int_{\nu_j}^{~\nu_j+1}{J_{\nu,j}\sigma_i(\nu)\nu^{-1}d\nu}},
\end{equation}
while the radiative recombination rate is simply given by 
\begin{equation}
R_{i+1,i}(T_e) = \alpha_{i+1,i} n_{i+1} n_e
\end{equation}
$\alpha_{i+1,i}$ here is the recombination rate coefficient from the ground state
of the upper ion to the ground state of the lower ion- we thus neglect recombination
to and from excited states in this calculation. In addition to these
radiative rates, we include dielectronic recombination and collisional ionization
as additive processes. We assume dilute level populations 
for simple-ions.

% We improve on this method by abandoning the modified Saha approach
% entirely, and instead computing the ionization state by 
% explicitly solving the rate equations in between ions in non-LTE. 
% Photoionization and recombination rates are calculated using MC estimators recorded
% during the photon propagation. Not only does this dispense with a number of
% small assumptions made in the modified Saha approach, it is also more numerically stable, 
% and in principle allows the direct addition of extra physical processes such as Auger ionization, which would have to be approximated if using the previous technique.

@@

We use the same atomic data as described by LK02 and since
updated by H13 and M15, with the addition of direct ionization data from Dere (xxxx). 
Photoionization cross-sections are from \textsc{Topbase} \citep{cunto1993} and  \cite{vfky}.
We use dielectronic recombination from \cite{badnell2006} and free-free 
Gaunt factors from \cite{sutherland1998}.

@@

\subsection{A Benchmark Model for BALQSOs}

\cite{higginbottom2013} presented a benchmark model for
(BAL)QSOs...introduce key parameters.

\subsubsection{Photon Sources}

@@

% \begin{table}
% \begin{tabular}{p{3cm}p{4cm}}
% \hline Free Parameters 	&	 Value \\ 
% \hline \hline 
% $M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
% $\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
% $\alpha_X$ 	 &	 $-0.9$ \\ 
% $L_{X} $ 	 &	 $1\times10^{43}~\rm{ergs~s^{-1}}$\\ 
% $r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
% $r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
% $\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
% $r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
% $r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
% $\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
% $\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
% $\lambda$ 	&	 $0$ \\ 
% $v_{\infty}$ 	&	 $v_{esc}$(f=1) \\ 
% $R_v$ 	        &	 $1\times10^{18}$cm \\ 
% $\alpha$ 	&	 $1.0$ \\
% \hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
%  $\alpha_{OX} $ 	 &	 -2.2\\ 
% \end{tabular}
% \caption{Wind geometry parameters used in the benchmark model.}
% \label{wind_param}
% \end{table}

\subsection{Potential for unification}

What was wrong with H13 model - X-rays + BELs.

@@

Introduce microclumping.

% To take account of clumping in our outflow we adopt a simple parameterization
% used in stellar wind modelling, known as {\em microclumping} \citep{hamann1998}. 
% The key assumption here is that typical clump sizes
% are much smaller than the typical photon mean free path, and thus the clumps are 
% both geometrically and optically thin. This approach is typically 
% known as microclumping and allows one to introduce a `filling factor', $f$, which is the 
% fraction of the volume of the plasma filled by clumps. We can then introduce the 
% density enhancement, $D$, which is simply

% \begin{equation}
% D = \frac{1}{f}
% \end{equation}

% One can then multiply all densities in the model by $D$, and all emitting volumes
% by $f$, meaning that all $\rho^2$ processes 
% (such as collisional excitation and recombination) will be enhanced, 
% while opacities remain unchanged (for fixed ionization state). A factor $f$ is also
% applied to the opacities such that opacities which scale only with $\rho$ are not
% increased. Clumping the wind has an important effect on the ionization state and has
% been proposed as a solution to the so-called `over-ionization problem' in 
% disc winds (refs). This is the main motivation for incorporating microclumping
% into our model.

@@

We arrived at a model which is one of the most promising,
but also representative of a family of models. Examination 
of this individual model allows us to gain insight
into fundamental geometrical and physical constraints, 
as discussed in section 5. 
The full grid, including output spectral files and plots can be found at
\url{jhmatthews.github.io/agn-grid-archive}.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/sightlines.eps}
\caption
{
The classes of sightlines and the corresponding possible 
system types in the framework of our model.
[FIGURE NEEDS IMPROVING]
}
\label{fig:lobal}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/xray_spectrum_lnu.eps}
\caption
{
Broadband SEDs compared to IR and X-ray SEDs for selected BALQSOs 
from Grupe \& Nousek (2015).
}
\label{fig:xray}
\end{figure}

@@

Although the low inclination, `BALQSO-like' viewing angles show moderate agreement with the data,
it appears that our models tend to underpredict the emergent X-ray luminosity at 
although we are limited by poor statistics. 
It is possible that this is due to our wind being overly optically thick to 
electron scattering at angles which look through the wind. It could also be that the
isotropic X-ray source assumption is incorrect and BALQSOs are {\em intrinsically} 
X-ray weak, as suggested by REFS. Nevertheless, our input X-ray spectrum
now reproduces the X-ray properties of a luminous quasar.

@@

We can also examine the emergent X-ray spectra from our model. 
Our code is not yet optimized for producing X-ray spectra, as we do not yet 
include all lines and levels of e.g. Fe or macro-atom treatments of 
higher order ions (cf. Sim et al. 2008, 2010). 
However, the bound-free opacity sources are complete, and comparisons of the shape
of the heavily absorbed spectra can be useful.
Figure 6 shows the broadband SED from the model, compared to a number of BALQSO spectra
from Grupe \& Nousek (2015) - note that lower mass and super-Eddington sources
are excluded.

@@

\subsection{LoBALs and trends with inclination}

@@

\clearpage
\section{Discussion}

@@

Can GR effects offer a potential solution? (not quite!)

{\bf Figures 8 and 9: AGNSPEC F(2000) as a function of viewing angle compared to a BB disc.
spectrum compared to composites with AGNSPEC correction.}

@@

A brief comment, citing Goodrich / Krolik \& Voit on the 
way in which anisotropy / attenuation affect the inferred
BAL fraction. We also need to be aware that there will be a number of selection
effects in building up the composites, and we should discuss these
and the subtleties involved.

>>>>

%% Journal definitions
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\apjs{ApJs}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}
\def\aaps{A\&As}

@@

\title
{
Modelling quasar outflows:
clumpy winds and broad emission lines
}

@@

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is normally attributed to emission from a geometrically thin, 
optically thick accretion disc surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material, 
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution (REFs) 
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

Approximately $20\%$ of quasars exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.
In addition, ADWs may offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) is normally
assumed to correspond to the dense wind base.
There are various spectroscopic sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit higher
ionization line absorption; LoBALQSOs which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

% As well as acting as a source of photoionized plasma, a 
As well as imprinting clear line absorption features,
disc wind may also have a profound effect on the structure and 
emergent {\em continuum} of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN \citep{laordavis2014}.
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN spectral energy distributions (SEDs) can, in general, be fitted well with accretion disc models.
Mass-loss therefore appears to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

@@

The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma. \cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.
An alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \cite{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required in line-driven hot star winds 
in order to explain observations \citep{hillier1991eswingsmodel}. Complex substructures 
on a variety of scales are also produced in simulations of line-driven 
outflows in AGN \citep{PSK2000,PK04,progakurosawa2010,proga2014}.
Clumpy winds therefore offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

@@

There has been some success using simple
kinematic prescriptions for biconical disc winds to model 
AGN and quasar outflows \citep{simlong2008,higginbottom2013}. 
\cite[][hereafter H13]{higginbottom2013},
successfully produced a benchmark model for BALQSOs. However, the model
had two key drawbacks. Firstly, an unrealistically low X-ray luminosity
was required in order to prevent over-ionization of the outflow.
Secondly, the model failed to produce the strong emission lines 
required at low inclinations in a unified model.
In this paper, we attempt to address these issues, and test the disc wind 
unification model using Monte Carlo radiative transfer (MCRT) and photoionization 
calculations. The paper is organised as follows.
In section 2, we describe our code. In section 3, we outline the model, including 
a description of our clumping prescription. In section 4, we present the results 
from a clumped model. In section 5 we discuss our results, 
focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interacting with either `simple ions'
or `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
The scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions. In this study,
only H \& He are treated as macro-atoms, because 
we expect recombination to be important
in determining their level populations (see M15).

@@

To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
we model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,j}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $j$, or an exponential 
\begin{equation}
J_{\nu,j}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $i$ can then be written as 
\begin{equation}
R_{i,i+1}(J)= 
\displaystyle{
n_i \left(C_{i} n_e + 
\sum_{band~j=0}^{n}~
{\int_{\nu_j}^{~\nu_j+1}{J_{\nu,j}\sigma_i(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_i$ is the photoionization cross-section and $C_{i}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $i$ is simply given by 
\begin{equation}
R_{i+1,i}(T_e) = (\alpha^i_{RR} + \alpha^i_{DR} + \alpha^i_{CR}) n_{i+1} n_e
\end{equation}
Where each $\alpha$ here is the recombination rate coefficient into the ground state of ion $i$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination respectively.
We thus neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

% We improve on this method by abandoning the modified Saha approach
% entirely, and instead computing the ionization state by 
% explicitly solving the rate equations in between ions in non-LTE. 
% Photoionization and recombination rates are calculated using MC estimators recorded
% during the photon propagation. Not only does this dispense with a number of
% small assumptions made in the modified Saha approach, it is also more numerically stable, 
% and in principle allows the direct addition of extra physical processes such as Auger ionization, which would have to be approximated if using the previous technique.
\subsection{Physical Processes}

We include all free-free, bound-free and bound-bound heating
and cooling processes in the model. For radiative transfer purposes
we treat electron scattering in the Thomson limit, 
but take full account of Compton heating and cooling when
calculating the thermal balance of the plasma (see H13).
Adiabatic cooling is included and represents the only
departure from strict radiative equilibrium, but is insignificant 
in most of the outflow.
It is dealt with self consistently by the spontaneous destruction
of $k$-packets with a probability

\begin{equation}
P_{\rm{adiabatic~destruction}} = \frac{C_{\rm{adiabatic}}}{C_{\rm{total}}},
\end{equation}
where $C_{\rm{adiabatic}}$ and ${C_{\rm{total}}}$ represent the 
adiabatic and total cooling rates in the cell.

@@

We use the same atomic data as described by LK02 and since updated by H13 and M15, 
with the addition of direct ionization data from \cite{dere2007}. 
Photoionization cross-sections are from \textsc{Topbase} \citep{cunto1993} and  \cite{vfky}.
Dielectronic and radiative recombination rate coefficients are taken from 
the \textsc{Chianti} database version 7.0 \citep{dere1997,landi2012}.
We use ground state recombination rates from \cite{badnell2006} where available,
and otherwise default to calculating recombination rates from the Milne
relation. Free-free Gaunt factors are from \cite{sutherland1998}.

@@

% \subsection{A Benchmark Model for BALQSOs}

% \cite{higginbottom2013} presented a benchmark model for
% (BAL)QSOs...introduce key parameters.

\subsection{Photon Sources}

@@

\subsection{Kinematics and Geometry}

Describe briefly the kinematics and geometry of the model.

@@

To take account of clumping in our outflow we adopt a simple parameterization
used in stellar wind modelling, known as {\em microclumping} \citep{hamann1998}. 
The key assumption here is that typical clump sizes
are much smaller than the typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach is typically 
known as microclumping and allows one to introduce a `filling factor', $f$, which is the 
fraction of the volume of the plasma filled by clumps. We can then introduce the 
density enhancement, $D$, which is simply defined as 
\begin{equation}
D = \frac{1}{f}.
\end{equation}
We then multiply all densities in the model by $D$, and all emitting volumes
by $f$, meaning that all $\rho^2$ emissivities and opacities
(such as collisional excitation and recombination) will be enhanced, 
while all $\rho$ process (such as electron scattering and bound-free opacity)
remain unchanged for a given ionization state. 
% A factor $f$ is also
% applied to the opacities such that opacities which scale only with $\rho$ are not
% increased.

Clumping the wind has an important effect on the ionization state and has
been proposed as a solution to the so-called `over-ionization problem' in 
disc winds (REFs). This is the main motivation for incorporating microclumping
into our model. This treatment is first-order; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless, clumping is clearly
important in these flows, and this parameterization allows a simple estimate
for the effect clumping might have on the ionization state and emergent 
line emission. It is also encouraging that microclumping has been used 
successfully in fits to O-star wind spectra (REFs).

@@

We arrived at a model which is one of the most promising,
but also representative of a family of models. Examination 
of this individual model allows us to gain insight
into fundamental geometrical and physical constraints, 
as discussed in section 5. 
The full grid, including output spectral files and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/sightlines.eps}
\caption
{
The classes of sightlines and the corresponding possible 
system types in the framework of our model.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:lobal}
\end{figure}

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/sed_all_balqsos.eps}
\caption
{
Broadband SEDs compared to IR and X-ray SEDs for selected BALQSOs 
from Grupe \& Nousek (2015).
}
\label{fig:xray}
\end{figure} %fullpage

@@

Although the low inclination, `BALQSO-like' viewing angles show moderate agreement with the data,
it appears that our models tend to underpredict the emergent X-ray luminosity at 
although we are limited by poor statistics. 
It is possible that this is due to our wind being overly optically thick to 
electron scattering at angles which look through the wind. It could also be that the
isotropic X-ray source assumption is incorrect and BALQSOs are {\em intrinsically} 
X-ray weak \citep[e.g.][]{morabito2013}. Nevertheless, our input X-ray spectrum
now reproduces the X-ray properties of a luminous quasar.

@@

We can also examine the emergent X-ray spectra from our model. 
Our code is not yet optimized for producing X-ray spectra, as we do not yet 
include all lines and levels of e.g. Fe or macro-atom treatments of 
higher order ions (cf. Sim et al. 2008, 2010). 
However, the bound-free opacity sources are complete, and comparisons of the shape
of the heavily absorbed spectra can be useful.
Figure 6 shows the broadband SED from the model, compared to a number of BALQSO spectra
from \cite{grupenousek2015} - note that lower mass and super-Eddington sources
are excluded.

@@

\subsection{LoBALs and ionization stratification}

@@

\clearpage
\section{Discussion: The Disc SED}

@@

General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}.
To generate GR disc spectra, we use the code \agn\ \citep{hubeny2000,davishubeny2006,davis2007}. 
The output flux at $2000$~\AA\ as a function of inclination
for an \agn\ model with the same disc and BH parameters
as our clumpy wind model is shown in figure~\ref{f2000}.
The effect of relativistic effects on disc anisotropy can be clearly seen.

To examine the effect of GR on our models, we carry out the following procedure.
First, we fit our output spectrum with a polynomial. This continuum fit can 
be subtracted from the actual spectrum 
to produce a model of the line emission only. We then calculate the 
`wind continuum factor', by comparing the model fit with the emergent 
disc spectrum without a wind. This factor gives us an approximate way to quantify the 
effect of the wind on a disc continuum. It will generally be $>1$ at high inclinations,
and $<1$ at low inclinations. This calculates emergent accretion disc spectra from an AGN disc atmosphere
including relativistic effects. Finally, produce our final synthetic spectrum 
by convolving this the line emission spectrum with the \agn\ model for the model parameters.

Figure 10 shows the resultant spectrum. Including GR clearly helps

% Can GR effects offer a potential solution? (not quite!)

% {\bf Figures 8 and 9: AGNSPEC F(2000) as a function of viewing angle compared to a BB disc.
% spectrum compared to composites with AGNSPEC correction.}

@@

The disc SED has a profound effect on the intrinsic BAL fraction
inferred from flux-limited samples. This in turn impacts on 
the geometry and covering factor assumed for BALQSO and unified wind models,
and even effects the efficiency of feedback.
\cite{krolik1998} noted that the effect of limb darkening and foreshortening
would result in a signicantly underestimated covering factor for BALQSO winds.
The effect is complicated further by the quasar luminosity function -- if 
BALQSOs have lower observed luminosities for given accretion disc properties 
then they may be drawn from a different bin in the quasar luminosity function 
\citep{goodrich1997}. These effects are understandably not taken into account
by current estimates of the intrinsic BALQSO fraction \citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
% These factors will also combine in the building-up 
% of BALQSO composite spectra, and mean that is very difficult to 
% disentangle the true nature of BALQSOs
% without BH mass and eddington ratio estimates.
The way forward is somewhat unclear. 
Given the lack of available observational constraints, models cannot readily 
be used to inform the selection criteria when calculating the BALQSO fraction.
Nevertheless, understanding the true nature of the disc SED, including the
angular distribution of radiation, is clearly crucial in order
to accurately estimate the true covering factor of BALQSO winds.

% A brief comment, citing Goodrich / Krolik \& Voit on the 
% way in which anisotropy / attenuation affect the inferred
% BAL fraction. We also need to be aware that there will be a number of selection
% effects in building up the composites, and we should discuss these
% and the subtleties involved.
